---
title: "eknma - Simulated Data Analysis"
format: html
editor: visual
---

# ECT v Ket NMA data simulation - Analysis and Plots

Using data simulated from aggregate values to conduct analysis

```{r}
library(dplyr)
library(tidyr)
library(purrr)
library(readxl)
library(here)
```

```{r}
simdat <- read_excel(here("data","processed","2026-04-22-eknma-simdata.xlsx"))
glimpse(simdat)
```

Set vars as factors

```{r}
simdat <- simdat %>%
  mutate(across(c(study, modality, treatment), as.factor))
str(simdat)
```

I want to simplify the "treatment" variable (e.g. remove dose details)

```{r}
library(stringr)

simdat <- simdat %>%
  mutate(
    dose = str_extract(treatment, "\\d+\\s*mg"),
    treatment = str_trim(str_remove(treatment, "[-\\s]*\\d+\\s*mg"))
  )

#for placbo, treatment should be either saline or midazolam
simdat <- simdat %>%
  mutate(treatment = if_else(treatment == "placebo", "saline", treatment), #if treatment = placebo change to saline, otherwhise keep original values 
         treatment = if_else(treatment == "ketamine", "racemic", treatment),
         treatment = as.factor(treatment))
#check levels
levels(simdat$treatment)

```

```{r}
#note. different variations of same treatment (e.g. esket 56mg vs 84mg) are treated as separate arms
arm_summary_detail <- simdat %>%
  group_by(study, modality, treatment, arm.name, timepoint, week) %>%   
  summarise(
    n = n(),
    mean = mean(y, na.rm = TRUE),
    sd = sd(y, na.rm = TRUE),
    .groups = "drop"
  )

#version now where they are collapsed
arm_sum_collapsed <- simdat %>%
  group_by(study, week, modality, trt = treatment, arm.name) %>%
  summarise(
    mean = mean(y, na.rm = TRUE),
    sd   = sd(y, na.rm = TRUE),
    n    = dplyr::n(),
    .groups = "drop"
  ) %>%
  distinct(study, week, trt, .keep_all = TRUE)
```

# Summary of data availability
## By timepoint
```{r}
simdat %>%
  group_by(week, treatment) %>%
  summarise(
    n_studies = n_distinct(study),
    n_participants = n(),
    .groups = "drop"
  ) %>%
  arrange(week, treatment)
```
```{r}
simdat %>%
  group_by(study, week, treatment, modality) %>%
  summarise(
    n_arms = n_distinct(arm.name),
    n_modality = n_distinct(modality),
    n_treatments = n_distinct(treatment),
    n_participants = n(),
    .groups = "drop"
  ) %>%
  arrange(study, week)
```
```{r}
simdat %>%
  group_by(study, week, modality,treatment) %>%
  summarise(n = n(), .groups = "drop") %>%
  pivot_wider(
    names_from = week,
    values_from = n
  ) %>%
  select(study, modality,"0","1","2","3","4")
```

# Plotting

```{r}
library(ggplot2)
```

```{r}
ggplot(data = arm_summary, aes(x = week, y = mean)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(aes(colour = study),
             stroke = 1.5,
             size = 3) +
  geom_line(aes(colour = study), size = 1) + 
  geom_errorbar(aes(ymin = mean - (sd/sqrt(n)), 
                    ymax = mean + (sd/sqrt(n)), 
                    colour = study), 
                size = 1,
                width = 0.1)
```

Testing breaking the data into ECT and ketamine subsets

```{r}
summary.ect <- arm_summary |>
  filter(modality == "ECT")
summary.ket <- arm_summary |>
  filter(modality %in% c("esketamine", "ketamine"))
```

```{r}
ggplot(arm_summary, 
       aes(x = week,
           y = mean,
           group = interaction(study, treatment))) +
  geom_point(size = 3) +
  geom_line(aes(colour = treatment)) +
  labs(x = "Timepoint", y = "Mean outcome") +
  theme_minimal()
```

I am confused because different studies have different comparisons...not sure what to do
Move on:

#arm level network plots
Making network plots this way means i don't need to make contrasts/pairwise comparisons
```{r}
library(dplyr)
library(purrr)
```

## By treatment
```{r}
edges_trt <- arm_summary %>%
  filter(week == 4) %>%
  group_by(study) %>%
  group_split() %>%
  map_dfr(function(d) {

    trts <- unique(d$treatment)

    # drop studies with only one treatment
    if (length(trts) < 2) return(NULL)

    comb <- t(combn(trts, 2))

    data.frame(
      from = comb[,1],
      to   = comb[,2]
    )
  }) %>%
  count(from, to, name = "weight") #add edge weights

nodes_trt <- arm_summary %>%
  filter(week == 4) %>%
  group_by(treatment) %>%
  summarise(size = sum(n), .groups = "drop") %>%
  rename(name = treatment)
```

```{r}
library(igraph)
library(ggraph)
```
```{r}
g_trt <-graph_from_data_frame(edges_trt, vertices = nodes_trt)
ggraph(g_trt, layout = "auto") +
  geom_edge_link(aes(width = weight), colour = "grey60") +
  geom_node_point(aes(size = size), colour = "steelblue") +
  geom_node_text(aes(label = name), repel = TRUE) +
  scale_edge_width(range = c(0.5, 2)) +
  scale_size_continuous(range = c(3, 10)) +
  theme_void()
```
## By modality
```{r}
edges_modal <- arm_summary %>%
  filter(week == 4) %>%
  group_by(study) %>%
  group_split() %>%
  map_dfr(function(d) {

    trts <- unique(d$modality)

    # drop studies with only one treatment
    if (length(trts) < 2) return(NULL)

    comb <- t(combn(trts, 2))

    data.frame(
      from = comb[,1],
      to   = comb[,2]
    )
  }) %>%
  count(from, to, name = "weight") #add edge weights

nodes_modal <- arm_summary %>%
  filter(week == 4) %>%
  group_by(modality) %>%
  summarise(size = sum(n), .groups = "drop") %>%
  rename(name = modality)

g_modal <- graph_from_data_frame(edges_modal, vertices = nodes_modal)
ggraph(g_modal, layout = "auto") +
  geom_edge_link(aes(width = weight), colour = "grey60") +
  geom_node_point(aes(size = size), colour = "steelblue") +
  geom_node_text(aes(label = name), repel = TRUE) +
  scale_edge_width(range = c(0.5, 2)) +
  scale_size_continuous(range = c(3, 10)) +
  theme_void()
```

## by treatment but with placebo combined
```{r}
edges_placebo <- arm_summary %>%
  mutate(
    trt_clean = case_when(
      str_detect(str_to_lower(treatment), "saline|midazolam") ~ "placebo",
      TRUE ~ treatment
    )
  ) %>%
  filter(week == 4) %>%
  group_by(study) %>%
  group_split() %>%
  map_dfr(function(d) {

    trts <- unique(d$trt_clean)

    # drop studies with only one treatment
    if (length(trts) < 2) return(NULL)

    comb <- t(combn(trts, 2))

    data.frame(
      from = comb[,1],
      to   = comb[,2]
    )
  }) %>%
  count(from, to, name = "weight")
nodes_placebo <- arm_summary %>%
  mutate(
    trt_clean = case_when(
      str_detect(str_to_lower(treatment), "saline|midazolam") ~ "placebo",
      TRUE ~ treatment
    )
  ) %>%
  filter(week == 4) %>%
  group_by(trt_clean) %>%
  summarise(size = sum(n), .groups = "drop") %>%
  rename(name = trt_clean)

g_placebo <- graph_from_data_frame(edges_placebo, vertices = nodes_placebo)

ggraph(g_placebo, layout = "fr") +
  geom_edge_link(aes(width = weight), colour = "grey60") +
  geom_node_point(aes(size = size), colour = "steelblue") +
  geom_node_text(aes(label = name), repel = TRUE) +
  scale_edge_width(range = c(0.5, 2)) +
  scale_size_continuous(range = c(3, 10)) +
  theme_void()
```

# Pairwise comparisons
## by modality  e.g. ketamine, esketamine, placebo, ECT -> note, this excludes Loo2015 PW study
```{r}
library(netmeta)
#pairwise by MODALITY e.g. ketamine, esketamine, placebo, ECT -> note, this excludes Loo2015 PW study
pw_modality <- arm_sum_collapsed  %>%
  group_split(week) %>%
  lapply(function(d) {

    # remove studies with only one treatment (e.g. ECT vs ECT after collapsing)
    d <- d %>%
      group_by(study) %>%
      filter(n_distinct(modality) > 1) %>%
      ungroup()

    # skip if nothing left
    if (nrow(d) == 0) return(NULL)

    pw <- pairwise(
      treat = d$modality,
      mean  = d$mean,
      sd    = d$sd,
      n     = d$n,
      studlab = d$study,
      sm = "MD"
    )

    pw$week <- unique(d$week)
    pw
  }) %>%
  bind_rows()
```

## By treatment
```{r}
pw_trt <- arm_sum_collapsed %>% 
  group_split(week) %>% 
  lapply(function(d) { 
    pairwise( treat = d$trt, mean = d$mean, sd = d$sd, n = d$n, studlab = d$study, sm = "MD" ) %>% 
      mutate(week = unique(d$week)) }) %>% 
  bind_rows()
```

## combining placebo methods
```{r}
pw_trt_collapsed <- simdat %>%
  mutate(
    trt_clean = case_when(
      str_detect(str_to_lower(treatment), "saline|midazolam") ~ "placebo",
      TRUE ~ treatment
    )
  ) %>%
  group_by(study, week, trt_clean) %>%
  summarise(
    n = n(),
    mean = mean(y, na.rm = TRUE),
    sd = sd(y, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  group_split(week) %>%
  map_dfr(function(d) {

    # drop studies that collapse to single treatment
    d <- d %>%
      group_by(study) %>%
      filter(n_distinct(trt_clean) > 1) %>%
      ungroup()

    if (nrow(d) == 0) return(NULL)

    pw <- pairwise(
      treat = d$trt_clean,
      mean  = d$mean,
      sd    = d$sd,
      n     = d$n,
      studlab = d$study,
      sm = "MD"
    )

    pw$week <- unique(d$week)
    pw
  })
```


# NMA
Start with modelling NMAs at 4 week timepoint
First treatment with placebo collapsed
```{r}
#by treatment, placebo collapsed
net_w4_trtcollapsed <- netmeta(
  TE = TE,
  seTE = seTE,
  treat1 = treat1,
  treat2 = treat2,
  studlab = studlab,
  data = pw_trt_collapsed %>% filter(week == 4),
  sm = "MD", random = TRUE, reference.group = "placebo"
)
netgraph(net_w4_trtcollapsed)
forest(net_w4_trtcollapsed)
```
Second: modality -> note Loo2015 excluded from analysisis. Only ECT data is ekstrand and andand
```{r}
net_w4_modality <- netmeta(
  TE = TE,
  seTE = seTE,
  treat1 = treat1,
  treat2 = treat2,
  studlab = studlab,
  data = pw_modality %>% filter(week == 4),
  sm = "MD", random = TRUE, reference.group = "placebo"
)
netgraph(net_w4_modality)
forest(net_w4_modality)
```
Obviously this is not a very strong analysis because of the poor network and the fact that all ECT vs placebo evidence is indirect

### Extrapolate NMA longitudinally

#### By treatment with placebo collapsed
```{r}
nma_by_week_trtcollapsed <- pw_trt_collapsed %>%
  split(.$week) %>%
  map(function(d) {
    netmeta(
      TE = d$TE,
      seTE = d$seTE,
      treat1 = d$treat1,
      treat2 = d$treat2,
      studlab = d$studlab,
      sm = "MD",
      random = TRUE,
      reference.group = "placebo"
    )
  })
```

```{r}

#set treatments except placebo
all_trts <- unique(unlist(map(nma_by_week_trtcollapsed, ~ .x$trts)))
all_trts <- setdiff(all_trts, "placebo")

#extract effects over time
results_trtcollapsed <- map_df(all_trts, function(trt) {
  map_df(names(nma_by_week_trtcollapsed), function(w) {
    net <- nma_by_week_trtcollapsed[[w]]
    trts <- net$trts
    if (!all(c(trt, "placebo") %in% trts)) {
      return(data.frame(
        week = as.numeric(w),
        treatment = trt,
        effect = NA,
        se = NA
      ))
    }

    data.frame(
      week = as.numeric(w),
      treatment = trt,
      effect = net$TE.random[trt, "placebo"],
      se = net$seTE.random[trt, "placebo"]
    )
  })
})

```

```{r}
ggplot(results_trtcollapsed,aes(x = week, y = effect, colour = treatment)) +
  geom_line() +
  geom_point() +
  labs(
    y = "Mean difference vs placebo",
    x = "Week"
  ) +
  theme_minimal()
```
### Making a function
```{r}
run_nma_by_week <- function(data,
                            treatments = NULL,
                            comparator = "placebo",
                            sm = "MD",
                            random = TRUE) {
  # Step 1: run NMA by week
  nma_list <- data %>%
    split(.$week) %>%
    map(~ netmeta(
      TE = .x$TE,
      seTE = .x$seTE,
      treat1 = .x$treat1,
      treat2 = .x$treat2,
      studlab = .x$studlab,
      sm = sm,
      random = random,
      reference.group = comparator
    ))

  # Step 2: auto-detect treatments if not provided
  if (is.null(treatments)) {
    treatments <- unique(unlist(map(nma_list, ~ .x$trts)))
    treatments <- setdiff(treatments, comparator)
  }

  # Step 3: extract results
  results <- map_df(treatments, function(trt) {
    map_df(names(nma_list), function(w) {

      net <- nma_list[[w]]
      trts <- net$trts

      if (!all(c(trt, comparator) %in% trts)) {
        return(data.frame(
          week = as.numeric(w),
          treatment = trt,
          effect = NA,
          se = NA
        ))
      }

      data.frame(
        week = as.numeric(w),
        treatment = trt,
        effect = net$TE.random[trt, comparator],
        se = net$seTE.random[trt, comparator]
      )
    })
  })

  return(results)
}


```

### by modality
```{r}
results_modality <- run_nma_by_week(data = pw_modality, comparator = "placebo")
results_modality
ggplot(results_modality,aes(x = week, y = effect, colour = treatment)) +
  geom_line() +
  geom_point() +
  labs(
    y = "Mean difference vs placebo",
    x = "Week",
    color = "Treatment"
  ) +
  theme_minimal()
```


Notable difference at baseline

```{r}
simdat.wbl <- simdat %>%
  group_by(participant_id) %>%
  mutate(
    baseline = y[week == 0][1]
  ) %>%
  ungroup()
```










