---
title: "ECTvKet NMA 2026.04.16"
format: html
editor: visual
---

# Simulating data for Ket v ECT IPD

Studies included:

-   Ekstrand 2022 - KetECT study
-   Fedgchin 2019 - esketamine study
-   Loo 2015 - RUL pulse width study
-   Loo 2023 - KADS data
-   Ochs-Ross 2020
-   Popova 2019
-   Takahashi 2021
-   Chen 2023
-   Fu 2020
-   Ionescu 2021
-   Martin 2020 - BL pulse width study \<- decided to exclude this because of small numbers
-   TREK (simulated parameters)

Second step:

-   simulate TREK data
    -   anchor week 4 to EOT
    -   kads + 2 mean and use same SDs - esketamine should be same at baseline
        -   how much higher the means are in esketamine is SD x effect size / divide by each week
    -   0.6 difference between the two arms -esketamine higher scores (worse)
-   need: Shiroma means and SDs

## Set up

### Packages

```{r}
library(dplyr)
library(tidyr)
library(purrr)
library(MASS)
library(readxl)
```

### Get data

```{r}
dat <- read_excel("data/raw/2026-04-16-ECTvKet_NMA_sim_data.xlsx")
```

```{r}
str(dat)

```

convert grouping vars into factors

```{r}
dat <- dat %>%
  mutate(across(c(study, modality, treatment), as.factor))
str(dat)
```

Question for the future - should year be treated numerically/as integer?

## Create Function

Set up correlation matrices:

Exchangeable - fixed correlation between all timepoints for an individual (user provides rho)

```{r}
make_exchangeable_cor <- function(k, rho) {
  R <- matrix(rho, nrow = k, ncol = k)
  diag(R) <- 1
  R
}

```

Ar(1) - ar1 style correlation matrix where timepoints closer together are more closely correlated

```{r}
make_ar1_cor_week <- function(week, rho) {
  outer(week, week, function(i, j) rho^abs(i - j))
} #week is var used to order timepoints
```

```{r}
simulate_longitudinal <- function(df_group,
                                              rho = 0.5,
                                              cor_structure = c("exchangeable", "ar1")) {
  cor_structure <- match.arg(cor_structure)

  if (!is.numeric(rho) || length(rho) != 1 || !is.finite(rho) || abs(rho) >= 1) {
    stop("rho must be a single finite number with abs(rho) < 1")
  }

  df_group <- df_group %>%
    dplyr::mutate(
      cont.n = as.numeric(cont.n),
      cont.mean = as.numeric(cont.mean),
      cont.sd = as.numeric(cont.sd),
      week = as.numeric(week)
    ) %>%
    dplyr::filter(
      !is.na(timepoint),
      !is.na(week),
      !is.na(cont.n),
      cont.n > 0,
      !is.na(cont.mean),
      !is.na(cont.sd)
    ) %>%
    dplyr::arrange(week)

  k <- nrow(df_group)

  if (k == 0) {
    return(tibble::tibble(
      timepoint = character(),
      week = numeric(),
      participant_id = integer(),
      y = numeric()
    ))
  }

  n_baseline <- as.integer(df_group$cont.n[1])

  if (!is.finite(n_baseline) || n_baseline <= 0) {
    return(tibble::tibble(
      timepoint = character(),
      week = numeric(),
      participant_id = integer(),
      y = numeric()
    ))
  }

  R <- switch(
    cor_structure,
    exchangeable = make_exchangeable_cor(k, rho),
    ar1 = make_ar1_cor_week(df_group$week, rho)
  )

  Z <- MASS::mvrnorm(
    n = n_baseline,
    mu = rep(0, k),
    Sigma = R
  )

  Y <- matrix(NA_real_, nrow = n_baseline, ncol = k)

  for (j in seq_len(k)) {
    target_mean <- df_group$cont.mean[j]
    target_sd   <- df_group$cont.sd[j]

    if (!is.finite(target_mean) || !is.finite(target_sd)) {
      Y[, j] <- NA_real_
    } else if (target_sd == 0) {
      Y[, j] <- target_mean
    } else {
      z <- Z[, j]
      z <- (z - mean(z)) / stats::sd(z)
      Y[, j] <- target_mean + target_sd * z
    }
  }

  out <- tibble::as_tibble(Y)
  names(out) <- paste0("t", seq_len(k))

  out <- out %>%
    dplyr::mutate(participant_id = seq_len(n_baseline)) %>%
    tidyr::pivot_longer(
      cols = tidyselect::starts_with("t"),
      names_to = "time_index",
      values_to = "y"
    ) %>%
    dplyr::mutate(time_index = as.integer(sub("t", "", time_index))) %>%
    dplyr::transmute(
      timepoint = df_group$timepoint[time_index],
      week = df_group$week[time_index],
      participant_id = participant_id,
      y = y
    )

  out
}

simulate_longitudinal_dataset <- function(dat,
                                                rho = 0.5,
                                                cor_structure = c("exchangeable", "ar1")) {
  cor_structure <- match.arg(cor_structure)

  dat %>%
    dplyr::group_by(study, modality, treatment, arm.name, scale) %>%
    dplyr::group_modify(~simulate_longitudinal(
      df_group = .x,
      rho = rho,
      cor_structure = cor_structure
    )) %>%
    dplyr::ungroup()
}
```

## Run function

```{R}
#rho set as 0.8
sim_data1 <- simulate_longitudinal_dataset(
  dat,
  rho = 0.8,
  cor_structure = "ar1"
)
```

check output

```{r}
dplyr::glimpse(sim_data1)
```

n's for all timepoints should reflect baseline ns

```{r}
sim_data1 %>%
  dplyr::count(study, arm.name, timepoint)
```

create nested df grouped by study

```{r}
#tidyverse method
study_nested1 <- sim_data1 %>%   
  dplyr::group_by(study) %>%   
  tidyr::nest()
```

```{r}
#base R
study_list1 <- split(sim_data1, sim_data1$study)
```

## Checking Output

Checking how simulated means/SDs compare to provided dat df

```{r}
#reloading because MASS packages interferes
library(dplyr)
library(tidyr)

check_data <- sim_data1 %>%
  dplyr::group_by(study, modality, scale, timepoint) %>%
  dplyr::summarise(
    sim_n = sum(!is.na(y)),
    sim_mean = mean(y, na.rm = TRUE),
    sim_sd = stats::sd(y, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::left_join(
    dat,
    by = c("study", "modality", "scale", "timepoint")
  )

check_data

```

```{r}
wide_dat1 <- sim_data1 %>%
  tidyr::pivot_wider(
    id_cols = c(study, modality, treatment, arm.name, scale, participant_id),
    names_from = week,
    values_from = y
  ) %>% {
    # identify week columns
    week_cols <- setdiff(names(.), c("study", "modality", "treatment", "arm.name", "scale", "participant_id"))
    
    # sort them numerically
    week_cols_sorted <- sort(as.numeric(week_cols))
    
    # reselect in correct order
    dplyr::select(., study, modality, treatment, arm.name, scale, participant_id,
                  all_of(as.character(week_cols_sorted)))
  }
  
```

```{r}
#check correlation matrix
wide_dat1 %>%
  dplyr::select(-study, -modality, -treatment, -arm.name, -scale, -participant_id) %>%
  cor(use = "pairwise.complete.obs")

```

Correlation betwen timepoint 4 and 0 = 0.264 i.e. quite low. Lets increase rho

```{r}
#rho set as 0.90
sim_data2 <- simulate_longitudinal_dataset(
  dat,
  rho = 0.90,
  cor_structure = "ar1"
)
```

```{r}
wide_dat2 <- sim_data2 %>%
  tidyr::pivot_wider(
    id_cols = c(study, modality, treatment, arm.name, scale, participant_id),
    names_from = week,
    values_from = y
  ) %>% {
    # identify week columns
    week_cols <- setdiff(names(.), c("study", "modality", "treatment", "arm.name", "scale", "participant_id"))
    
    # sort them numerically
    week_cols_sorted <- sort(as.numeric(week_cols))
    
    # reselect in correct order
    dplyr::select(., study, modality, treatment, arm.name, scale, participant_id,
                  all_of(as.character(week_cols_sorted)))
  }
```

```{r}
#check correlation matrix
wide_dat2 %>%
  dplyr::select(-study, -modality, -treatment, -arm.name, -scale, -participant_id) %>%
  cor(use = "pairwise.complete.obs")
```

Thats better, correlation between baseline and week 4 timepoints is \~0.496

```{r}
writexl::write_xlsx(sim_data2,"data/processed/2026-04-22-eknma-simdata.xlsx")
```
