---
title: "ECTvKet NMA 2026.04.16"
format: html
editor: visual
---

# Simulating data for Ket v ECT IPD

Studies included:

-   Ekstrand 2022 - KetECT study
-   Fedgchin 2019 - esketamine study
-   Loo 2015 - RUL pulse width study
-   Loo 2023 - KADS data
-   Martin 2020 - BL pulse width study \<- decided to exclude this because of small numbers

Second step:

-   simulate TREK data
    -   anchor week 4 to EOT
    -   kads + 2 mean and use same SDs - esketamine should be same at baseline
        -   how much higher the means are in esketamine is SD x effect size / divide by each week
    -   0.6 difference between the two arms -esketamine higher scores (worse)
-   need: Shiroma means and SDs

## Set up

### Packages

```{r}
library(dplyr)
library(tidyr)
library(purrr)
library(MASS)
library(readxl)
```

### Get data

```{r}
dat <- read_excel("data/ECTvKet-NMA.simdata2026.04.16.xlsx")
```

```{r}
str(dat)
```

sd is character - fix this (note. this is now built into the sim_data function, no need to run)

```{r}
dat <- dat %>%
  mutate(
    cont.sd = as.numeric(cont.sd),
    across(c(study, modality, treatment), as.factor))
str(dat)
```

# Create a function to simulate data

2 ways of thinking about time - discrete and growth curves (days)

## Correlation matrices

matrix where all sets of timepoints have the same correlation

```{r}
make_exchangeable_cor <- function(k, rho) {
  R <- matrix(rho, nrow = k, ncol = k)
  diag(R) <- 1
  R
}
```

matrix nearby timepoints are more correlated - note. to use this would need to order the timepoints var sequentially (currently inconsistent)

```{r}
make_ar1_cor <- function(k, rho) {
  outer(seq_len(k), seq_len(k), function(i, j) rho^abs(i - j))
}
```

## Function

```{r}
simulate_longitudinal_group_exact <- function(df_group, #dataset assuming on row per arm and timepoint
                                              rho = 0.5,
                                              cor_structure = c("exchangeable", "ar1")) {
  cor_structure <- match.arg(cor_structure) #ensures that cor_structure is one of the allowed values
  #clean and prepares the input data
  df_group <- df_group %>% 
    dplyr::mutate(
      cont.n = as.numeric(cont.n),
      cont.mean = as.numeric(cont.mean),
      cont.sd = as.numeric(cont.sd)
    ) %>%
    #remove bad rows
    dplyr::filter( 
      !is.na(timepoint),
      !is.na(cont.n),
      cont.n > 0,
      !is.na(cont.mean),
      !is.na(cont.sd)
    ) %>%
    dplyr::arrange(timepoint)

  k <- nrow(df_group) #counts number of timepoints
#returns empty output if no rows remain after filtering bad rows
  if (k == 0) {
    return(tibble::tibble(
      timepoint = character(),
      participant_id = integer(),
      y = numeric()
    )) 
  }
#for a given study/arm, takes the max sample size across all timepoints because function simulates a full set of participants first, then introduces missingness at later timepoints)
  n_max <- max(df_group$cont.n) 
#safety check for where nmax is invalid
  if (!is.finite(n_max) || n_max <= 0) {
    return(tibble::tibble(
      timepoint = character(),
      participant_id = integer(),
      y = numeric()
    ))
  } 

  n_max <- as.integer(n_max)
#builds correlation matrix
  R <- switch(
    cor_structure,
    exchangeable = make_exchangeable_cor(k, rho),
    ar1 = make_ar1_cor(k, rho)
  ) 
#Simulate correlated standard normal data
  Z <- MASS::mvrnorm(n = n_max, mu = rep(0, k), Sigma = R) # z = a matric of n_max x k(timepoints)
  Y <- Z
#missingness matrix
  missing_mat <- matrix(FALSE, nrow = n_max, ncol = k)

  for (j in seq_len(k)) {
    n_obs <- as.integer(df_group$cont.n[j]) #get n for each timepoint
#randomly select subjects to be missing
    if (!is.na(n_obs) && n_obs < n_max) {
      miss_idx <- sample(seq_len(n_max), size = n_max - n_obs, replace = FALSE)
      missing_mat[miss_idx, j] <- TRUE
    }
  }
#rescale timepoint to match target mean and SD
  for (j in seq_len(k)) {
    target_mean <- df_group$cont.mean[j]
    target_sd   <- df_group$cont.sd[j]
    obs_idx <- which(!missing_mat[, j])

    if (length(obs_idx) == 0 || !is.finite(target_mean) || !is.finite(target_sd)) {
      Y[, j] <- NA_real_
    } else if (length(obs_idx) == 1 || target_sd == 0) {
      Y[, j] <- NA_real_
      Y[obs_idx, j] <- target_mean
    } else {
      z <- Z[obs_idx, j]
      z <- (z - mean(z)) / stats::sd(z)
      y_obs <- target_mean + target_sd * z

      Y[, j] <- NA_real_
      Y[obs_idx, j] <- y_obs
    }
  }
#convert simulated matrix to a tibble
  out <- tibble::as_tibble(Y)
  names(out) <- paste0("t", seq_len(k))
#reshape and add labels
  out <- out %>%
    dplyr::mutate(participant_id = seq_len(n_max)) %>%
    tidyr::pivot_longer(
      cols = tidyselect::starts_with("t"),
      names_to = "time_index",
      values_to = "y"
    ) %>%
    dplyr::mutate(time_index = as.integer(sub("t", "", time_index))) %>%
    dplyr::left_join(
      df_group %>% dplyr::mutate(time_index = seq_len(dplyr::n())),
      by = "time_index"
    )

  out[, c("timepoint", "participant_id", "y")]
}
```

## Wrapper

```{r}
simulate_longitudinal_dataset_exact <- function(dat, rho = 0.5,
                                                cor_structure = c("exchangeable", "ar1")) {
  cor_structure <- match.arg(cor_structure)

  dat %>%
    dplyr::group_by(study, modality, treatment, arm.name, scale) %>%
    dplyr::group_modify(~simulate_longitudinal_group_exact(
      df_group = .x,
      rho = rho,
      cor_structure = cor_structure
    )) %>%
    dplyr::ungroup()
}
```

## Run function

```{r}
set.seed(123)
sim_long <- simulate_longitudinal_dataset_exact(dat, rho = 0.5)
str(sim_long)
```

## Checking output

```{r}
check_long <- sim_long %>%
  dplyr::group_by(study, modality, scale, timepoint) %>%
  dplyr::summarise(
    sim_n = sum(!is.na(y)),
    sim_mean = mean(y, na.rm = TRUE),
    sim_sd = stats::sd(y, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::left_join(
    dat,
    by = c("study", "modality", "scale", "timepoint")
  )

check_long
```

# Simulating TREK data

Based on steves work: TREK.sample.simulations.r

## set parameters

```{r}
MEAN.DIFF <- 0
VAR <- 10
MEAN.KET <- 10
MEAN.ESKET <- 13

PRIOR.MU <- 5.64
PRIOR.SEM <- 6.6
```

## get data

```{r}
cov.data <- read_excel("data/KADS.bookend.xlsx")
```

```{r}
cov.data <- as.data.frame(cov.data)
cov.data <- cov.data[complete.cases(cov.data$MADRS_Total_ENDRCT), ]
cov(cov.data[,4:5])
```

```{r}
N <- 150
BASE.SD <- 6.5
END.SD <- 10
```

```{r}

matrix_data <- c(BASE.SD^2, (BASE.SD*END.SD)/2, 
                 (BASE.SD*END.SD)/2, END.SD^2)
cv_mat <- matrix(data = matrix_data, nrow = 2, ncol = 2, byrow = TRUE)
rownames(cv_mat) <- c("base","end")
colnames(cv_mat) <- c("base","end")
```

```{r}
#racemic
mns_1 <- c(32, 32 - 16) 
sim_data_rac <- round(mvrnorm(n = N, mu = mns_1, Sigma = cv_mat)); 
colnames(sim_data_rac) <- c("Baseline", "EndRCT"); 
subject <- 1:N 
cond <- rep(x = "racemic", times = N) 
sp_1 <- data.frame(subject, cond, sim_data_rac);
# ESKET
mns_2 <- c(32, 32 - 14)
sim_data_esk  <- round(mvrnorm(n = N, mu = mns_2, Sigma = cv_mat));
colnames(sim_data_esk) <- c("Baseline", "EndRCT");
subject <- 1:N
cond<- rep(x = "esketamine", times = N)
sp_2<- data.frame(subject, cond, sim_data_esk);

# Bind together into complete dataset
trek.sim <- rbind(sp_1, sp_2);
trek.sim$cond <- as.factor(trek.sim$cond)
str(trek.sim)
```

# Combining datasets

## Reshape simulated trek data

```{r}
trek_long <- trek.sim %>%
  rename( #to match sim_long data set
    participant_id = subject,
    modality = cond
  ) %>%
  pivot_longer(
    cols = c(Baseline, EndRCT),
    names_to = "timepoint",
    values_to = "y"
  ) %>%
  mutate( #to match sim_long data
    timepoint = recode(timepoint,
      Baseline = "pre",
      EndRCT = "EOT"
    ),
    study = "TREK",
    scale = "MADRS",
    treatment = modality,
    arm.name = modality
  ) 
```

## Make full dataset

```{r}
sim_long.all <- bind_rows(trek_long, sim_long)
```

```{r}
sim_long.all <- sim_long.all %>%
  dplyr::select(study, 
         modality, 
         treatment, 
         arm.name, 
         scale, 
         timepoint, 
         participant_id, 
         y)
```

```{r}
#tidyverse method
study_nested <- sim_long.all %>%   
  dplyr::group_by(study) %>%   
  tidyr::nest()
```

```{r}
#base R
study_list <- split(sim_long.all, sim_long.all$study)
```

```{r}
library(writexl)
write_xlsx(sim_long.all, "eknma.simdata.output.xlsx")
```

# Take 2
