---
title: "Ketamine MRFF"
author: "Nikolin"
date: "22/04/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# MRFF Study

## 1 Design

The primary outcome will be the MADRS score, a 10-item Likert scale with total scores ranging from 0 to 60. The endpoint is 4-weeks post-randomisation. There will be 2 treatment arms: Subcutaneous Racemic Ketamine, and Intranasal Spravato Esketamine. The comparison will be baseline-adjusted MADRS scores at 28 days after randomisation.

Of interest are:

-   showing that racemic ketamine is superior to esketamine
-   showing that racemic ketamine is non-inferior
-   showed equivence between racemic ketamine and esketamine

## 2 Outcome

The MADRS will be treated as a continuous outcome. A lower score indicates better outcomes. The KADS Cohort 2 group showed a reduction of 8 at the primary end-point (4-weeks) from a baseline of 30. Pooling largest esketamine trials (Ionescu et al., 2021; Fedgchin et al., 2019; Fu et al., 2020; Popova et al., 2019) showed a mean reduction of 24 at the primary end-point (3-4 weeks) from a baseline of 39.

For the present study, the mean value of the MADRS at baseline is expected to be 30, in line with KADS data from both cohorts and conditions. Given participants will be aware that they are receiving an active intervention (i.e., no placebo arm), we hypothesise a decrease in MADRS scores of 16 in the Subcutaneous Racemic Ketamine and group, approximately double that of the KADS cohort 2 improvement, and an SD of 12 in MADRS scores across groups.

## 3 Network Meta-Analysis

To estimate the indirect comparison of racemic ketamine vs esketamine, use a Bayesian network meta-analysis of 13 RCTs of repeated sessions of ketamine for the treatment of depression., with 1808 participants in 32 experiment arms.

The models predict a standardised mean difference (SMD) of 0.63 (95% CrI: -0.03, 1.31), which equates to a MADRS score difference of 5.65 (95% CrI: 1.63, 11.60) in favour of racemic ketamine.

```{r nma.section, echo=FALSE, results=FALSE}
library(esc)
library(gemtc)
library(multinma)
library(ggplot2)
library(stringr)
library(rjags)

# set labels
NEG_LABEL = "Favours intervention"
POS_LABEL = "Favours placebo"

# set file names
SMD.DATA <- 'C:\\Users\\Admin\\OneDrive - UNSW\\Studies\\Ketamine\\TBD - Ketamine TREK\\data\\df_smd.csv'

RAW.DATA <- 'C:\\Users\\Admin\\OneDrive - UNSW\\Studies\\Ketamine\\TBD - Ketamine TREK\\data\\df_raw.csv'

# load the data from a csv file
df_smd <- read.csv(SMD.DATA, header=TRUE)
df_raw <- read.csv(RAW.DATA, header=TRUE)

# plot a network based on a contrast-based effect size data frame
plot_network_with_multinma <- function(df, title){
  
  # create a multinma network object
  network = set_agd_contrast(df,
                             study = study,
                             trt = treatment,
                             y = diff,
                             se = std.err,
                             sample_size = n)
  
  # plot network; for style options see:
  # https://dmphillippo.github.io/multinma/reference/plot.nma_data.html
  plot(network, weight_edges=TRUE, weight_nodes=TRUE) +
    ggplot2::theme(legend.position = "bottom",
                   legend.box = "vertical",
                   legend.margin = ggplot2::margin(0,0,0,0),
                   legend.spacing = ggplot2::unit(0.5, "lines")) +
    ggplot2::ggtitle(title)
}

# plot the network diagram
plot_network_with_multinma(df_smd, "Antidepressant efficacy")
plot_network_with_multinma(df_raw, "Antidepressant efficacy")

# create a gemtc network object from the .csv data
smd.network = mtc.network(data.re = df_smd,
                          description = "Bayesian NMA data - SMD")
raw.network = mtc.network(data.re = df_raw,
                          description = "Bayesian NMA data - raw")

# create a random effects, normal likelihood, identity link model
smd.model = mtc.model(smd.network,
                      likelihood = "normal",
                      link = "identity",
                      linearModel = 'random')
raw.model = mtc.model(raw.network,
                      likelihood = "normal",
                      link = "identity",
                      linearModel = 'random')

# run MCMC - burn-in 50,000, iterations 100,000, thin 20
smd.mcmc1 = mtc.run(smd.model,
               n.adapt=50000,
               n.iter=100000,
               thin=20)
raw.mcmc1 = mtc.run(raw.model,
               n.adapt=50000,
               n.iter=100000,
               thin=20)

# forest plots relative to high-dose racemic ketamine
forest(relative.effect(smd.mcmc1, t1="Ket_High"), 
       digits=3,
       draw.no.effect=TRUE,
       right.label=POS_LABEL,
       left.label=NEG_LABEL,
       use.description=TRUE)

forest(relative.effect(raw.mcmc1, t1="Ket_High"), 
       digits=3,
       draw.no.effect=TRUE,
       right.label=POS_LABEL,
       left.label=NEG_LABEL,
       use.description=TRUE)

# SMD of MADRS only studies Ket_High vs Esket_High is 0.62
# MADRS_diff of MADRS only studies is 5.61
# Therefore, the SMD is 5.61/0.62 = 9.05

```

## 4 Approximate Fixed Sample Size

Assuming the following:

-   two-sided independent samples t-test
-   comparative effect size from NMA analysis, d = 0.63
-   $\alpha$ = 0.05
-   1 - $\beta$ = 95%

```{r simple.power.analysis, echo=FALSE}
library(pwr)
p.out <- pwr.t.test(d = 0.63, 
                    power = 0.90, 
                    sig.level = 0.05, 
                    type="two.sample")
plot(p.out)

```


## 5 Non-inferiority

Non-inferiority analyses using difference > -4

```{r simple.noninferiority.analysis, echo=FALSE}
library(TOSTER) 

# worst case scenario - no difference, large SD from ket C2 KADS
power_t_TOST(
  #n = 66,
  delta = 3,
  sd = 10,
  low_eqbound = -3,
  high_eqbound = Inf,
  alpha = 0.025,
  power = 0.90,
  type = "two.sample"
)

# 63 per group for 15%
# 67 per group for 10%

# or 


# NMA otucomes 
#   delta = 5.65
#   sd = 3.375
#   alpha = 0.025
#   power = 0.95
# Non-inf at -4.40 gives 80% power with n = 67
# Non-inf at -4.19 gives 80% power with n = 74 (added 10% from dropout margin)

# non-inf power, assuming 67 or 71 per group
ES = 0.63
ES = 0.3

power_t_TOST(
  n = NULL,
  delta = 0,
  sd = 10,
  low_eqbound = -4,
  high_eqbound = Inf,
  alpha = 0.025,
  power = 0.90,
  type = "two.sample"
)

```


## 6 Bayesian Modelling - Informative Priors

Permutations change the seed by one for each permutation produce a random distribution of 120 participants shift the full distribution by the 95% credible interval range (-2:12) subsample out of the full distribution (n.subset = 50:10:120)

```{r bayesian.model, echo=FALSE}
library(truncnorm)
library(brms) 

#library(tidyverse)  suppressWarnings(library(tidyverse))
library(ggridges) 
library(shinystan) 
library(bayesplot) 
library(tidybayes) 
library(ggmcmc)

# parameters
PERMUTATIONS <- 200
SAMPLE.RANGE <- seq(from = 60, to = 120, by = 10)
EFFECT.RANGE <- seq(from = -2, to = 12, by = 2)

# empty data frames
columns = c("madrs", "group", "treatment")
ket.data = data.frame(matrix(nrow=0, ncol =length(columns)))
esket.data = data.frame(matrix(nrow=0, ncol =length(columns)))

ket.data <- data.frame(madrs = numeric(),
                 group = factor(),
                 treatment = factor(),
                 stringsAsFactors = FALSE)

colnames(ket.data) = columns
colnames(esket.data) = columns

# assumed values based on largest trials
max.n <- 120
madrs.change.min <- -40
madrs.change.max <- 15
madrs.change.mean <- -16
madrs.change.sd <- 12.4

# set model priors
modelpriors1 = c(set_prior("normal(5.65,3.3075)", class = "b", coef = "groupesketamine"),
                 set_prior("normal(0,12.4)", class = "sigma"))

for (perm in 1:PERMUTATIONS) {
  for (madrs.diff in EFFECT.RANGE) {
    
    # ketamine distribution
    set.seed(perm)
    ket.distribution <- round(rtruncnorm(max.n,
                                         a = madrs.change.min,
                                         b = madrs.change.max,
                                         mean = madrs.change.mean,
                                         sd = madrs.change.sd))
    
    # esketamine distribution
    set.seed(PERMUTATIONS + perm)
    esket.distribution <- round(rtruncnorm(max.n,
                                         a = madrs.change.min,
                                         b = madrs.change.max,
                                         mean = madrs.change.mean + madrs.diff,
                                         sd = madrs.change.sd))
    
    for (samp in SAMPLE.RANGE) {
      
      # select subsets of the data above
      ket.subset <- sample(ket.distribution,samp) 
      esket.subset <- sample(esket.distribution,samp) 
      
      # create data frames for analysis
      ket.data <- data.frame(madrs = ket.subset,
                             group = factor(rep("ketamine",length(ket.subset))),
                             treatment = rep(0,length(ket.subset)))
      
      esket.data <- data.frame(madrs = esket.subset,
                             group = factor(rep("esketamine",length(esket.subset))),
                             treatment = rep(1,length(esket.subset)))
      
      # bind the data frames together
      pseudo.data <- rbind(ket.data, esket.data)

      
      # << COULD ADD IF STATEMENT TO UPDATE MODEL WITH LARGER SAMPLE >>
      
      if (samp > SAMPLE.RANGE[1]){
        
        # update model fit to save time
        updated.fit.model <- update(fit.model, 
                              newdata = pseudo.data, 
                              seed = perm*samp*(madrs.diff+abs(min(EFFECT.RANGE))+1)) 
        
        #   extract answers from research question hypotheses
        hyp.noninf <- hypothesis(updated.fit.model, "groupesketamine > -4")
        bayes.factor.noninf <- hyp.noninf$hypothesis$Evid.Ratio
        
        hyp.super <- hypothesis(updated.fit.model, "groupesketamine > 0")
        bayes.factor.super <- hyp.super$hypothesis$Evid.Ratio

      } else {
        
        # run bayesian fit - NEED TO SOLVE THE ISSUE OF INFORMATIVE PRIORS
        fit.model <- brm(madrs ~ 0 + Intercept + group,
                         family = gaussian(),
                         prior = modelpriors1,
                         seed = perm,
                         iter = 2000,
                         chains = 4,
                         warmup = 1000,
                         control = list(adapt_delta = 0.99),
                         data = pseudo.data)
        
        #   extract answers from research question hypotheses
        hyp.noninf <- hypothesis(fit.model, "groupesketamine > -4")
        bayes.factor.noninf <- hyp.noninf$hypothesis$Evid.Ratio
        
        hyp.super <- hypothesis(fit.model, "groupesketamine > 0")
        bayes.factor.super <- hyp.super$hypothesis$Evid.Ratio
        
      }

      
      # NOT NEEDED
      # outcomes of the model
      plot(fit.model) 
      print(fit.model) 
      fixef(fit.model)

      # extract bayes factor for comparison
      posterior_summary(fit.model)
      
      # store data
      #   the groupesketamine difference
      #   the bayesfactor for calculating
      
    }
  }
}
  

```

List of web-pages to look at:

<https://discourse.mc-stan.org/t/help-understanding-and-setting-informative-priors-in-brms/9574>

<https://rmarkdown.rstudio.com/articles_intro.html>

<https://rmarkdown.rstudio.com/authoring_basics.html>

<https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf>

<https://www.rstudio.com/blog/rstudio-v1-4-preview-visual-markdown-editing/>

<https://rpruim.github.io/s341/S19/from-class/MathinRmd.html>

<https://solomonkurz.netlify.app/post/bayesian-power-analysis-part-i/>

<https://marissabarlaz.github.io/portfolio/bayesian/#how-to-set-priors-in-brms>

<https://www.r-bloggers.com/2014/02/bayes-factor-t-tests-part-2-two-sample-tests/>

<https://www.r-bloggers.com/2019/05/bayesian-models-in-r-2/>

<file:///C:/Users/Admin/AppData/Local/Microsoft/Windows/INetCache/Content.Outlook/I4AXIAHQ/Sequential-Design.html>

<file:///C:/Users/Admin/AppData/Local/Microsoft/Windows/INetCache/Content.Outlook/I4AXIAHQ/20210701_design_notes.html>

<https://clayford.github.io/dwir/dwr_12_generating_data.html>

<https://stats.stackexchange.com/questions/30303/how-to-simulate-data-that-satisfy-specific-constraints-such-as-having-specific-m>

<https://stackoverflow.com/questions/32792824/create-a-fake-dataset-that-fits-the-following-parameters-n-mean-sd-min-and>

<https://bmcpsychiatry.biomedcentral.com/articles/10.1186/s12888-018-1761-4>

<https://cran.r-project.org/web/packages/pwr/vignettes/pwr-vignette.html>
